---
title: "RMarkDown"
author: "Krrish"
date: "March 4, 2018"
output:
  html_document:
    keep_md: true
---


##importing library
```{r , echo=TRUE}
library(plyr)
library(ggplot2)
library(lattice)
```
##reading csv
```{r, echo=TRUE}
health<-read.csv("activity.csv",na.strings = "NA",sep = ",",header = TRUE)
names(health)
summary(health)
head(health)
```
##converting date variable to date class
```{r, echo=TRUE}
health$date<-as.Date(health$date,format= "%Y-%m-%d")
health$interval<-factor(health$interval)
```
##ignore the missing case
```{r, echo=TRUE}
ignoreNa<- is.na(as.character(health$steps))
data_no_NA<- health[!ignoreNa,]
```

##aggregation the number of steps taken each day
```{r, echo=TRUE}
steps_each_day<-aggregate(steps~date,data =data_no_NA,sum )
head(steps_each_day)
```

## create histogram 
```{r, echo=TRUE}
hist(steps_each_day$steps,breaks = 20,col ="red",xlab = "number of steps",ylab="Date",
     main="histogram of total number of steps taken each day")
```
##Mean and median number of steps taken each day
```{r, echo=TRUE}
mean(steps_each_day$steps)

median(steps_each_day$steps)
```
##Time series plot of the average number of steps taken
```{r, echo=TRUE}
steps_per_interval<- aggregate(data_no_NA$steps,by=list(interval=data_no_NA$interval),FUN=mean)
head(steps_per_interval)
colnames(steps_per_interval)<-c("interval","average_steps")

plot(as.integer(levels(steps_per_interval$interval)),steps_per_interval$average_steps,type = "l",
     xlab = "Interval",ylab = "Average Number of Steps",main = "Average Daliy Activity pattern",col="blue")
```

####maximum number of average of steps
```{r, echo=TRUE}
maxNumberSteps<- max(steps_per_interval$average_steps)
maxNumberSteps
```

####The 5-minute interval that, on average, contains the maximum number of steps
```{r, echo=TRUE}
max_step<-steps_per_interval[which.max(steps_per_interval$average_steps),]$interval
max_step
```
##imputing missing value
```{r, echo=TRUE}
sum(is.na(as.character(health$steps)))
```
##for "date" variable
```{r, echo=TRUE}
sum(is.na(as.character(health$date)))
```
##for interval vlaue
```{r, echo=TRUE}
sum(is.na(as.character(health$interval)))
```
##histogram
```{r, echo=TRUE}
hist(as.numeric(steps_each_day$steps),breaks = 20,col="red", xlab = "Number of steps",main = "histogram of total number taken each day")
```

## create a factor variable to stor the day of week
```{r, echo=TRUE}
data_no_NA$day <- as.factor(weekdays(data_no_NA$date))
table(data_no_NA$day)
```
##creating coa logical variable for weekdays and weekend
```{r, echo=TRUE}
data_no_NA$is_weekday<- ifelse(!(data_no_NA$day %in% c("Saturday","Sunday")),TRUE,FALSE)
table(data_no_NA$is_weekday)
```

##calculatting average number of steps in week days
```{r, echo=TRUE}
weekdays_data<- data_no_NA[!data_no_NA$is_weekday,]
steps_per_interval_weekdays<-aggregate(weekdays_data$steps, by=list(interval=weekdays_data$interval),FUN=mean)
```
##calculating the average number of steps for weekends
```{r, echo=TRUE}
weekends_data<- data_no_NA[!data_no_NA$is_weekday,]
steps_per_interval_weekends <- aggregate(weekends_data$steps,by=list(interval=weekends_data$interval),FUN=mean)
```


##Adding columns names
```{r, echo=TRUE}
colnames(steps_per_interval_weekdays) <- c("interval", "average_steps")
colnames(steps_per_interval_weekends) <- c("interval", "average_steps")
```
##Adding a column to indecate the day
```{r, echo=TRUE}
steps_per_interval_weekdays$day <- "Weekday"
steps_per_interval_weekends$day <- "Weekend"
```
##merging the tow together
```{r, echo=TRUE}
week_data <-rbind(steps_per_interval_weekdays,steps_per_interval_weekends)
week_data$day <- as.factor(week_data$day)
```
##making plot
```{r, echo=TRUE}
xyplot(average_steps ~ interval|day,type="l",data = week_data,layout=c(1,2),
       ylab="Number of steps")

```
